---
title: "Supported Models and recipes steps"
description: >
  A complete list of supported parsnip models, recipes steps, and tailor
  adjustments that can be converted to orbital objects.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Supported Models and recipes steps}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The supported methods currently come from [tidypredict](https://tidypredict.tidymodels.org/) or are implemented natively for ANN models.

## Supported models

This table doesn't exhaustively list fully unsupported models. Please file [an issue](https://github.com/tidymodels/orbital/issues) to add model to table.

```{r}
#| echo: false
#| message: false
if (!rlang::is_installed(c("gt", "tibble"))) {
  knitr::knit_exit()
}
```

```{r}
#| echo: false
#| message: false
#| error: false
#| output: false
library(gt)
library(dplyr)

tibble::tribble(
  ~parsnip, ~engine, ~numeric, ~class, ~prob,
  "`boost_tree()`", "`\"catboost\"`", "✅", "✅", "✅",
  "`boost_tree()`", "`\"lightgbm\"`", "✅", "✅", "✅",
  "`boost_tree()`", "`\"xgboost\"`", "✅", "✅", "✅",
  "`cubist_rules()`", "`\"Cubist\"`", "✅", "❌", "❌",
  "`decision_tree()`", "`\"partykit\"`", "✅", "✅", "✅",
  "`decision_tree()`", "`\"rpart\"`", "✅", "✅", "✅",
  "`linear_reg()`", "`\"lm\"`", "✅", "❌", "❌",
  "`linear_reg()`", "`\"glmnet\"`", "✅", "❌", "❌",
  "`logistic_reg()`", "`\"glm\"`", "❌", "✅", "✅",
  "`logistic_reg()`", "`\"glmnet\"`", "❌", "✅", "✅",
  "`mars()`", "`\"earth\"`", "✅", "✅", "✅",
  "`multinom_reg()`", "`\"glmnet\"`", "❌", "✅", "✅",
  "`naive_Bayes()`", "`\"naivebayes\"`", "❌", "⚪", "⚪",
  "`nearest_neighbor()`", "`any`", "❌", "❌", "❌",
  "`rand_forest()`", "`\"randomForest\"`", "✅", "✅", "✅",
  "`rand_forest()`", "`\"ranger\"`", "✅", "✅", "✅",
  "`mlp()`", "`\"nnet\"`", "✅", "✅", "✅",
  "`mlp()`", "`\"brulee\"`", "✅", "✅", "✅",
  "`mlp()`", "`\"brulee_two_layer\"`", "✅", "✅", "✅",
  "`mlp()`", "`\"keras\"`", "✅", "✅", "✅",
  "`mlp()`", "`\"h2o\"`", "✅", "✅", "✅",
  "`mlp()`", "`\"kerasnip\"`", "✅", "✅", "✅"
) |>
  gt() |>
  tab_spanner(
    label = "Model",
    columns = c(parsnip, engine)
  ) |>
  tab_spanner(
    label = "Regression",
    columns = c(numeric)
  ) |>
  tab_spanner(
    label = "Classification",
    columns = c(class, prob)
  ) |>
  tab_header(
    title = "Supported Prediction Types"
  ) |>
  cols_align(
    "center",
    columns = c(numeric, class, prob)
  ) |>
  tab_footnote("✅: Supported") |>
  tab_footnote("❌: Cannot be supported") |>
  tab_footnote("⚪: Not yet supported") |>
  fmt_markdown(
    columns = c(parsnip, engine)
  )
```

## Artificial Neural Networks (MLP models)

orbital translates `mlp()` models from nnet, brulee, keras, keras3, kerasnip, and H2O
to SQL by unrolling every weight matrix and activation function into a chain of
SQL column expressions.

### Supported activations

All standard activations from PyTorch / brulee are supported, including `relu`,
`tanh`, `sigmoid`, `elu`, `selu`, `gelu`, `leaky_relu`, `softplus`, `silu`/`swish`,
and `mish`. See `orbital:::activation_expr()` for the full list.

### Probability output

For binary and multiclass MLP models orbital produces both class predictions and
per-class probability columns when you pass `type = "prob"` (or
`type = c("class", "prob")`):

```r
orb <- orbital(fit, type = "prob")
predict(orb, new_data)
#> # A tibble: 32 × 2
#>    .pred_0 .pred_1
#>      <dbl>   <dbl>
#>  1   0.621   0.379
#>  ...
```

For multiclass softmax models the probability expressions include the
`exp(logit) / sum(exp(logits))` normalisation.

### Keras Functional API (skip connections)

Sequential and linear Functional API models work out of the box.
Functional models that use `Add` layers (residual / skip connections) are
also supported — orbital detects the merge layers, parses the model config,
and generates the appropriate element-wise addition expressions.

```r
# Residual model — works
inp <- keras3::keras_input(shape = 4)
x   <- keras3::layer_dense(inp, 4, activation = "relu")
out <- keras3::layer_add(list(x, inp)) |>   # residual
          keras3::layer_dense(1)
model <- keras3::keras_model(inp, out)
orbital(model, mode = "regression", feature_names = paste0("x", 1:4))
```

### Caveats and recommendations

**SQL expression length** Deep networks (6+ hidden layers, 100+ units per layer)
generate very long SQL strings.  Some databases limit query length or query-plan
depth.  For very deep networks, consider materializing intermediate layer outputs
into temporary tables rather than executing one monolithic query.

**Runtime dependencies** keras and H2O models require a live Python / JVM
process to *fit* the model.  orbital only handles the *prediction* translation;
no Python/JVM process is needed at inference time once the orbital object is
created.

**H2O Maxout** The Maxout activation uses a grouped max-pooling architecture
that orbital does not yet support.  A call to `orbital()` on an H2O model trained
with `activation = "Maxout"` raises an informative error.  Use `Rectifier`,
`Tanh`, or `Sigmoid` for orbital-compatible H2O models.

## Recipes steps

```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
#| label: setup
#| echo: false
library(orbital)
```

```{r}
#| echo: false
all_funs <- ls(getNamespace("orbital"))

steps <- grep("orbital.step_", all_funs, value = TRUE)
steps <- gsub("orbital.", "", steps)
```

The following `r length(steps)` recipes steps are supported

```{r}
#| results: asis
#| echo: false
cat(paste0("- `", steps, "()`\n"))
```

## tailor adjustments

The following 4 tailor methods are supported


- `tailor::adjust_equivocal_zone()`
- `tailor::adjust_numeric_range()`
- `tailor::adjust_predictions_custom()`
- `tailor::adjust_probability_threshold()`
